### 2013-09-18 chi2014/results.tex
Some revision, fixing of lots of technical problems

Figure FIG shows the cumulative count of new ideas as a function of the number of instances gathered in the experiment. For reference, a perfect ``1:1'' idea generation line is also plotted.

@@

We fit the parameters to the data using a Bayesian approach [REF Kruschke], using the STAN modeling software to perform the actual fit [REF STAN]. In modeling the parameters, we use uninformative priors. The resulting posterior distributions allow us to determine whether the conditions are significantly different from one another by determining whether the highest density intervals (HDIs) of $b_{cond}$'s posterior distributions overlap or not. In our analyses, we use an HDI of 95\%.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{ideas_over_time}
    \caption{Cumulative idea count}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\columnwidth]{rate_new_idea_over_time}
    \includegraphics[width=0.7\columnwidth]{rate_new_category_over_time}
    \includegraphics[width=0.7\columnwidth]{rate_new_ns_category_over_time}
    \caption{Rate of idea generation}
\end{figure}

@@

\subsubsection{Most Popular Ideas}
Our results suggest that individuals typically produce ideas from a common set idea categories early in the brainstorming process, then diverge and produce more original ideas. This phenomenon can be observed by examining the o-scores of ideas and idea categories in Figures X and Y, which show the distribution of o-scores for individual ideas and idea categories, as a function of response number, across all conditions.

@@

To more explicitly confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total). If our intuition is correct, these ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances. We introduce two Bernoulli random variables to model this. The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run ($P(\theta_f|top)$). The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} (>5) instances in a run ($P(\theta_l|top)$).

@@

\subsubsection{Hypothesis 4}
The steady increase in originality seen in the first 20 instances (Figure~\ref{fig:idea_oscore_order}, \ref{fig:cat_oscore_order}), provides evidence for \textbf{Hypothesis 4:} ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
With the point of change seen at the 20 point, we use this as our division point, fitting two each for idea and category o-scores: instances in the first 20 of runs, and instances in the second 20 runs.
In examining the data, it became clear that o-scores do not follow a remotely normal distribution. Instead, we fit a beta distributions to each data division. The results are in Table~\ref{tab:hyp4} and Figure~\ref{fig:idea_oscore_hyp4}.

@@

The idea o-score models are significantly different in the $\alpha$ parameter, with mean o-scores of 0.995 and 0.997 for the $\leq20$ and $\textgreater20$ conditions respectively. The category o-scores are significantly different in both parameters, with mean o-scores of 0.975 and 0.982.

Participants generated more original ideas after their first 20 instances than they did after. Not only does this support the hypothesis, it provides a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for less from more.

\subsection{Idea Generation Patterns}
Given the clear differences in characteristics of ideas as a function of response number, we now consider idea generation patterns observable from the data. In this section, we examine these patterns in the context of hypotheses 5 and 6. Specifically, these hypotheses draw upon the SIAM model and predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 5). Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 6).

To determine category switching, we examine whether an idea was assigned to the same category tree as the previous, or whether it was assigned to a different category tree. We also examine the timing data collected from the web app to examine the time to generate an idea as a function of whether it is in the same category as the previous idea.

Our data show clear evidence in support of both hypotheses. As an example, below is an example of a worker riffing on a single idea.

@@

\begin{table}
\begin{tabular}[h!]{r | l l l l l l l}
    \hline \hline \textbf{condition} & 5 & 10 & 20  \\ \hline \hline
    \% riffs & 13.65 & 19.28 & 25.17  \\
    source instances & 10.58 & 12.92 & 15.01 \\
    riffs per source & 1.29 & 1.49 & 1.68 \\
    median length of consecutive chain & 2 & 2 & 2 \\
    median distance to previous in chain & 1 & 2 & 2 \\
    median distance to first in chain & 2 & 3 & 5 \\ \hline \hline
    \textbf{condition} & 50 & 75 & 100 \\ \hline \hline
    \% riffs & 39.6 & 32.81 & 42.81 \\
    source instances & 19 & 17.19 & 18.13\\
    riffs per source & 2.08 & 1.91 & 2.36\\
    median length of consecutive chain & 2 & 2 & 2\\
    median distance to previous in chain & 7 & 10 & 5\\
    median distance to first in chain & 16.5 & 18 & 26.5 \\
    \end{tabular}
    \caption{Riffing statistics by condition}
\end{table}

Examining our data, we find that the higher the number of instances requested, the more likely participants are to use old ideas as inspiration for new ideas (i.e., ``riff'' on previous ideas). However, these riff sequences are typically quite small in length, across conditions.

@@

To test hypothesis 6, we now examine the time it takes to generate a response instance.

An examination of the distribution of data for time spent per instance suggests a log-normal distribution to the data. Using STAN, we fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{hyp5_comparison}
    \caption{Time spent to generate an instance}
\end{figure}

The resulting fit models are shown in Figure FIG, overlapping the log-space histograms for observed values. The mean for the within-category condition was 9.5 in log space (13.4 seconds, 95\% HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, 95\% HDI 10.2-10.3). The variances were 1.1 (3 seconds) and 1.2 (3.3 seconds) respectively. These difference were both sigificant. Within-category instance generation took on average 12.6 seconds less than between-categories. This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between conditions. Thus, we find support for hypothesis 6 and also show the applicability of the SIAM model to brainstorming in microtask marketplaces.

>>>>

Figure~\ref{fig:cumulative_ideas} shows the cumulative count of new ideas as a function of the number of instances gathered in the experiment. For reference, a perfect ``1:1'' idea generation line is also plotted.

@@

We fit the parameters to the data using a Bayesian approach \cite{kruschke_doing_201}, using the STAN modeling software to perform the actual fit \cite{stan-software:2013}. In modeling the parameters, we use uninformative priors. The resulting posterior distributions allow us to determine whether the conditions are significantly different from one another by determining whether the highest density intervals (HDIs) of $b_{cond}$'s posterior distributions overlap or not. In our analyses, we use an HDI of 95\%.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{cumulative_ideas}
    \caption{Cumulative number of novel ideas}
    \label{fig:cumulative_ideas}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{cumulative_categories}
    \caption{Cumulative number of novel idea categories}
    \label{fig:cumulative_categories}
\end{figure}

% \begin{figure}[h!]
%     \centering
%     \includegraphics[width=0.7\columnwidth]{rate_new_idea_over_time}
%     \includegraphics[width=0.7\columnwidth]{rate_new_category_over_time}
%     \includegraphics[width=0.7\columnwidth]{rate_new_ns_category_over_time}
%     \caption{Rate of idea generation}
% \end{figure}

@@

\subsubsection{Most Popular Ideas}
Our results suggest that individuals typically produce ideas from a common set idea categories early in the brainstorming process, then diverge and produce more original ideas. This phenomenon can be observed by examining the o-scores of ideas and idea categories in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order}, which show the distribution of o-scores for individual ideas and idea categories, as a function of response number, across all conditions.

@@

To more explicitly confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total). If our intuition is correct, these ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances. We introduce two Bernoulli random variables to model this. The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run ($P(\theta_f|top)$). The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} ($\textgreater5$) instances in a run ($P(\theta_l|\overline{top})$).

@@

\subsubsection{Originality}

The increase in o-score over time in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order}) also supports hypothesis 4, suggest by Parnes et al \cite{parnes_effects_1961}, that ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
We test a stronger guarantee implied by our data: that ideas after the first 20 have higher o-scores than the first 20.
Although we expected the o-scores to be normally distributed in each condition, they were not. Instead, we fit Beta distributions to o-score for each of the $\leq20$, $\textgreater20$ conditions. Table~\ref{tab:hyp4} and Figure~\ref{fig:idea_oscore_hyp4} show the resulting parameters and models.

@@

The $\alpha$ parameter is significantly different in both the idea and category o-score comparison. The mean idea o-score is 0.997 in the $\textgreater20$ versus 0.995 in the $\leq20$ condition. Similarly, the $\leq20$ condition is favourable in category o-score with a mean of 0.982 vs 0.975.

This support the hypothesis, it provides a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for less from more.

\subsection{Idea Generation Patterns}
Given the clear differences in characteristics of ideas as a function of response number, we now consider idea generation patterns observable from the data. In this section, we examine these patterns in the context of hypotheses 5 and 6. Specifically, these hypotheses draw upon the SIAM model's conceptualization of brainstorming as a process in ideas are generated within one category image at a time.

In our data, we identify this pattern of creating new instances based on old ones as ``riffing'' (Table~\ref{tab:riffing}). The higher the number of instances requested, the more likely participants were to use old ideas as inspiration for new ideas (i.e., ``riff'' on previous ideas). However, these riff sequences are typically quite small in length, across conditions.

@@

\begin{table}
\begin{tabular}[h!]{r | l l l l l l l}
    \hline \hline \textbf{condition} & 5 & 10 & 20  \\ \hline \hline
    \% riffs & 13.65 & 19.28 & 25.17  \\
    source instances & 10.58 & 12.92 & 15.01 \\
    riffs per source & 1.29 & 1.49 & 1.68 \\
    median length of consecutive chain & 2 & 2 & 2 \\
    median distance to previous in chain & 1 & 2 & 2 \\
    median distance to first in chain & 2 & 3 & 5 \\ \hline \hline
    \textbf{condition} & 50 & 75 & 100 \\ \hline \hline
    \% riffs & 39.6 & 32.81 & 42.81 \\
    source instances & 19 & 17.19 & 18.13\\
    riffs per source & 2.08 & 1.91 & 2.36\\
    median length of consecutive chain & 2 & 2 & 2\\
    median distance to previous in chain & 7 & 10 & 5\\
    median distance to first in chain & 16.5 & 18 & 26.5 \\
    \end{tabular}
    \caption{Riffing statistics by condition}
    \label{tab:riffing}
\end{table}

As an example of riffing, below is an excerpt of a particular participant's brainstorming run.

\begin{quote}
    \begin{enumerate}
        \item recycle
        \item take apart and refurbish
        \item scrap it for any precious metals inside (some have gold contacts)
    \end{enumerate}
\end{quote}

The participant first touches on the general idea of recycling. Then, they become more specific by reducing the scope of recyling to specific parts. Finally, they limit the scope of parts to precious metals, particularly gold. At this point, the participant exhausts their working memory of the category an moves on.

SIAM predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 5). Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 6).

\subsubsection{Category changing}

To determine category switching, we examine whether an idea was assigned to the same category tree as the previous, or whether it was assigned to a different category tree.

@@

\subsubsection{Idea generation time}

To test hypothesis 6, we now examine the time it takes to generate a response instance. We calculate this as the time from the first activation of the form widget to the last.

An examination of the distribution of data for time spent per instance suggests a log-normal distribution to the data. We fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{hyp5_comparison}
    \caption{Time spent to generate an instance}
    \label{fig:hyp5_comparison}
\end{figure}

The resulting fit models are shown in Figure~\ref{fig:hyp5_comparison}, overlapping the log-space histograms for observed values. The mean for the within-category condition was 9.5 in log space (13.4 seconds, 95\% HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, 95\% HDI 10.2-10.3). The variances were 1.1 (3 seconds) and 1.2 (3.3 seconds) respectively. These difference were both sigificant. Within-category instance generation took on average 12.6 seconds less than between-categories. This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between conditions. Thus, we find support for hypothesis 6 and also show the applicability of the SIAM model to brainstorming in microtask marketplaces.
